\hypertarget{recommendations}{\subsection{Recommendations}}
\label{sec:recommendations}

On the basis of our findings, we make specific recommendations aimed at foundation model developers, foundation model deployers, and policymakers in \refsec{recommendations}. 
We highlight our top-most recommendation for each stakeholder group.

\paragraph{Foundation model developers should improve transparency by drawing on the practices of their competitors.}
By assessing developers directly, we clarify for each developer the indicators where they lack transparency.
In itself, this provides a clear diagnostic on where they stand relative to their competitors and, given our justifications for why transparency on these matters is valuable, why improving transparency would be beneficial for society.
Given that \numfeasible indicators are all satisfied by some developer, developers can directly consult the practices of their competitors to provide a clear example of how they might improve their transparency.
There is a tremendous gap between the \numfeasible already-feasible indicators and the current top score of \maxscore and the mean score of \meanscore, meaning there are many areas of low-hanging fruit where developers can readily improve transparency today. 

\paragraph{Foundation model deployers should push for greater transparency from developers.}
Foundation models intermediate a growing supply chain: deployers of foundation models (\eg cloud service providers and companies that license developers' models) as well as other downstream actors are influenced by, and can influence, the transparency of foundation model developers.
In particular, deployers should push developers for greater transparency when making the decision, and potentially negotiating the contract, to deploy a developer's model.
Deployers and other downstream actors wield leverage collectively: it is their downstream use that generates users and revenue for foundation model developers, meaning they should use this leverage to acquire the necessary transparency from foundation model developers.

\paragraph{Policymakers should prioritize transparency with sufficient precision.}
Given the importance of transparency, policymakers should make transparency a top priority in legislative proposals and regulatory enforcement related to foundation models. 
While transparency is already broadly recognized in most regulatory frameworks for AI, policymakers should be more precise about what they mean by transparency and the areas in which they hope to reduce opacity via transparency requirements or other measures.
In particular, policymakers should understand the status quo for transparency (\eg via the scores we provide) and use this evidence to inform interventions in the areas where transparency is most urgently needed (\eg on \labor and \impact, given these are lowest-scoring dimensions of transparency across the entire supply chain). 
\clearpage