\paragraph{Acknowledgements.}
We thank Alex Engler, Anna Lee Nabors, Anna-Sophie Harling, Arvind Narayanan, Ashwin Ramaswami, Aspen Hopkins, Aviv Ovadya, Benedict Dellot, Christie Lawrence, Connor Dunlop, Conor Griffin, Dan Ho, Dan Jurafsky, Deb Raji, Dilara Soylu, Divyansh Kaushik, Gerard de Graaf, Iason Gabriel, Irene Solaiman, John Hewitt, Joslyn Barnhart, Judy Shen, Madhu Srikumar, Marietje Schaake, Markus Anderljung, Mehran Sahami, Neel Guha, Peter Cihon, Peter Henderson, Rebecca Finlay, Rob Reich, Rohan Taori, Rumman Chowdhury, Russell Wald, Seliem El-Sayed, Seth Lazar, Stella Biderman, Steven Cao, Tatsu Hashimoto, Toby Shevlane, Vanessa Parli, Yann Dubois, Yo Shavit, and Zak Rogoff for discussions on the topics of foundation models, transparency, and/or indexes that informed the \projectname. 
We especially thank Loredana Fattorini for her extensive work on the visuals for this project, as well as Shana Lynch for her work in publicizing this effort.

\paragraph{Foundation Model Developers.}
We thank the following individuals at their respective organizations for their engagement with our effort, including involvement in responding to our initial scores on behalf of their organizations. 
We emphasize that \textbf{this acknowledgement should not be understood as an endorsement of the \projectname by these individuals}, but simply that they were involved in our engagement with their organizations.
-  \aitwentyone. Yoav Shoham
-  \amazon. Bratin Saha, Vasi Philomin, Atul Deo, Swami Sivasubramanian, Peter Hallinan
-  \anthropic. Jack Clark, Deep Ganguli, Thomas Liao
-  \cohere. Aidan Gomez, Danielle Smalls, Seraphina Goldfarb-Tarrant, Nick Jakobi, Saurabh Baji
-  \google. Slav Petrov, James Manyika, Kremena Goranova, Sarah Portik, Alexandra Belias
-  \huggingface. Clement Delangue, Meg Mitchell, Yacine Jernite
-  \inflection. Mustafa Suleyman, Tim Hwang
-  \meta. Joelle Pineau, Melanie Kambadur, Joe Spisak, Eric Smith, Louis Martin
-  \openai. Miles Brundage, Lama Ahmad
-  \stability. Emad Mostaque, Ben Brooks

\paragraph{Funding.}
This work was supported in part by the 2022 Hoffman-Yee program at the Stanford Institute for Human-Centered Artificial Intelligence (HAI).\footnote{\url{https://hai.stanford.edu/2022-hoffman-yee-grant-recipients}}
This work was supported in part by the AI2050 program at Schmidt Futures (Grant G-22-63429).

\paragraph{Conflict of Interest.}
Given the nature of this work (\eg potential to significantly impact particular companies and shape public opinion), we proactively bring attention to any potential conflicts of interest, deliberately taking a more expansive view of conflict of interest to be especially forthcoming.

-  Betty Xiong is not, and has not, been affiliated with any of the companies evaluated in this effort or any other private sector entities.
-  Daniel Zhang is not, and has not, been affiliated with any of the companies evaluated in this effort or any other private sector entities.
-  Kevin Klyman is not, and has not, been affiliated with any of the companies evaluated in this effort or any other private sector entities.
-  Nestor Maslej is not, and has not, been affiliated with any of the companies evaluated in this effort or any other private sector entities.
-  Percy Liang was a post-doc at Google (September 2011--August 2012), a consultant at Microsoft (May 2018--May 2023), and a co-founder of Together AI (July 2022--present).
- He is not involved in any other companies.
-  Rishi Bommasani is not, and has not, been affiliated with any of the companies evaluated in this effort.
- Rishi is an author of \citet{jernite2022governance}, as part of the BigScience initiative, that guided the data governance practices for developing BLOOM.
- As a result, he is also an author on the 350+ author BLOOM paper \citep{scao2022bloom} that is often cited in the scoring of \bloomz.
-  Sayash Kapoor worked at Meta until December 2020. He has not since worked for the company.
-  Shayne Longpre has three connections to the assessed developers.
- He has worked as a Student Researcher at Google Brain in 2022, and is an on-going contributor to Cohere For AI, Cohere's non-profit volunteer research organization.
- Lastly, he was part of the BigScience initiative, where he contributed to BLOOM \citep{scao2022bloom}.