\hypertarget{author-contributions}{\section{Author contributions}}
\label{app:author-contributions}
This project was a team effort, built on countless contributions from everyone involved. 
Contributions span the conceptualization, design, execution, analysis, and presentation phases, along with the overarching vision, organization, coordination, and leadership.
Below, we describe each team member's contribution.


\noindent 
\textbf{Betty Xiong}: Betty contributed to the design, execution, analysis, and presentation.
\\
\textbf{Daniel Zhang}: Daniel contributed to the design, execution, and presentation.
\\
\textbf{Kevin Klyman}: Kevin co-led the project and coordinated nearly all aspects of the initiative. Kevin contributed to the design, execution, analysis, and presentation. 
\\
\textbf{Nestor Maslej}: Nestor contributed to the design, execution, analysis, and presentation.
\\
\textbf{Percy Liang}: Percy advised the project, shaping the initial vision. Percy contributed to the conceptualization, design, analysis, and presentation along with providing overarching guidance on all stages. 
\\
\textbf{Rishi Bommasani}: Rishi led the project, bringing together the team, providing the initial vision, and coordinating all aspects of the initiative. Rishi contributed to the conceptualization, design, execution, analysis, and presentation. 
\\
\textbf{Sayash Kapoor}: Sayash contributed to the design, execution, analysis, and presentation.
\\
\textbf{Shayne Longpre}: Shayne contributed to the design, execution, analysis, and presentation.
\\